---
title: "ARMA/ARIMA/SARIMA Models"
author: "Yifeng Chen"
date: "2023-04-26"
output: html_document
---

```{r, echo=FALSE, include=FALSE}
library(plotly)
library(dplyr)
library(gridExtra)
library(ggfortify)
library(vars)
library(GGally)
library(lmtest)
library(ggplot2)
library(tseries)
library(forecast)
```

```{r, echo=FALSE, include=FALSE}
grossDP_us <- read.csv("Gross Domestic Product US.csv")
consumerPI_us <- read.csv("Consumer Price Index US.csv")
realDI_us <- read.csv("Real Disposable Personal Income US.csv")
unemploymentR_us <- read.csv("Unemployment Rate US.csv")
```

```{r, echo=FALSE, include=FALSE}
grossDP_us$DATE <- as.Date(grossDP_us$DATE)
consumerPI_us$DATE <- as.Date(consumerPI_us$DATE)
realDI_us$DATE <- as.Date(realDI_us$DATE)
unemploymentR_us$DATE <- as.Date(unemploymentR_us$DATE)
```

```{r, echo=FALSE, include=FALSE}
colnames(grossDP_us)[2] <- "DATA"
colnames(consumerPI_us)[2] <- "DATA"
colnames(realDI_us)[2] <- "DATA"
colnames(unemploymentR_us)[2] <- "DATA"
```

```{r, echo=FALSE, include=FALSE}
consumerPI_us <- consumerPI_us %>%
  mutate(RATE = (DATA - lag(DATA)) / lag(DATA) * 100)
```

```{r, echo=FALSE, include=FALSE}
merged_data <- Reduce(function(x, y) merge(x, y, by = "DATE"), list(grossDP_us, consumerPI_us, realDI_us, unemploymentR_us))
colnames(merged_data)[2] <- "grossDP_us"
colnames(merged_data)[3] <- "consumerPI_us"
colnames(merged_data)[4] <- "consumerPIR_us"
colnames(merged_data)[5] <- "realDI_us"
colnames(merged_data)[6] <- "unemploymentR_us"
```

```{r, echo=FALSE, include=FALSE}
remove_missing_values <- function(data) {
  return(data[!is.na(data)])
}

# Consumer Price Index Rate
data_cpir <- ts(remove_missing_values(consumerPI_us$RATE), frequency = 12, start = c(1947, 2))

# Unemployment Rate
data_unemployment <- ts(remove_missing_values(unemploymentR_us$DATA), frequency = 12, start = c(1976, 1))

# Consumer Price Index
data_cpi <- ts(remove_missing_values(consumerPI_us$DATA), frequency = 12, start = c(1947, 1))

# Gross Domestic Product
data_gdp <- ts(remove_missing_values(grossDP_us$DATA), frequency = 12, start = c(1947, 1))

# Real Disposable Income
data_rdi <- ts(remove_missing_values(realDI_us$DATA), frequency = 12, start = c(1951, 1))
```


```{r, echo=FALSE}
# Consumer Price Index
adf_cpi <- adf.test(log(data_cpi))
print("ADF test for Consumer Price Index")
print(adf_cpi)

# Gross Domestic Product
adf_gdp <- adf.test(log(data_gdp))
print("ADF test for Gross Domestic Product")
print(adf_gdp)

# Real Disposable Income
adf_rdi <- adf.test(log(data_rdi))
print("ADF test for Real Disposable Income")
print(adf_rdi)
```

```{r, echo=FALSE}
diff1_data_cpi <- diff(data_cpi)

adf.test(diff1_data_cpi)
```

```{r, echo=FALSE}
diff1_data_gdp <- diff(data_gdp)

adf.test(diff1_data_gdp)
```

```{r, echo=FALSE}
diff1_data_rdi <- diff(data_rdi)

adf.test(diff1_data_rdi)
```

**By the first differencing now we can use all the variables to predict unemployment rate.**

**The main goal is to predict the unemployment rate, so for the actual AR/MA/ARMA/ARIMA goal for this section, we will use unemployment rate on that.**

```{r, echo=FALSE}
# Original data plot
orig_plot <- autoplot(data_unemployment) + ggtitle("Unemployment Rate in the U.S.")

# ACF plot for original data
acf_plot <- ggAcf(data_unemployment) + ggtitle("ACF of Unemployment Rate in the U.S.")

# PACF plot for original data
pacf_plot <- ggPacf(data_unemployment) + ggtitle("PACF of Unemployment Rate in the U.S.")

# Group plots together
grid.arrange(orig_plot, acf_plot, pacf_plot, ncol = 1)
```

```{r, echo=FALSE}
# Unemployment Rate
adf_unemployment <- adf.test(data_unemployment)
print("ADF test for Unemployment Rate")
print(adf_unemployment)
```

**From the unemployment rate plot and the ADF test we can see that we do not need to further difference the data**
**The ARIMA would be (1, 0, 0) and suggest AR(1)**

```{r, echo=FALSE}
# Fit ARIMA(1,0,1) model to time series data
arima_model <- arima(data_unemployment, order = c(1, 0, 0))
arima_model
```

**Equation: Y_t = c + ϕ_1 Y_{t-1} + ε_t**

```{r, echo=FALSE}
auto.arima(data_unemployment)
```

**Since the ARIMA(1, 0, 0) has no AICc or BIC because only the PACF gives a value out of boundary at 1 and the data is already stationary without further differencing, we would use AIC to compare the 2 and the result is that the ARIMA(1, 0, 0) has a lower AIC than the auto-arima one, so we would go with the orignal one.**

```{r, echo=FALSE}
checkresiduals(arima_model)
```

```{r, echo=FALSE}
# Extract residuals from the model
residuals <- residuals(arima_model)

# Create Q-Q plot of the residuals
qqnorm(residuals)
qqline(residuals)
```

```{r, echo=FALSE}
# Forecast future values with 95% confidence interval
forecast_result <- forecast(arima_model, h = 10, level = 0.95)

# Plot forecast values and confidence intervals
autoplot(forecast_result) + ggtitle("Forecasts from ARIMA (1, 0, 0)")
```

```{r}
# Split time series data into training and testing sets
train_data <- data_unemployment[1:720]
test_data <- data_unemployment[721:903]

# Fit ARIMA(1,0,0) model to training data
arima_model <- arima(train_data, order = c(1, 0, 0))

# Forecast future values with 95% confidence interval
forecast_result <- forecast(arima_model, h = length(test_data), level = 0.95)

# Calculate accuracy metrics for ARIMA model
accuracy_result_arima <- accuracy(forecast_result, test_data)

# Fit benchmark methods to training data
naive_model <- naive(train_data, h = length(test_data))

# Calculate accuracy metrics for benchmark methods
accuracy_result_naive <- accuracy(naive_model, test_data)

print("Accuracy metrics for ARIMA model:")
print(accuracy_result_arima)
print("Accuracy metrics for Naive method:")
print(accuracy_result_naive)
```

```{r}
# Combine the actual data, training data, and test data forecasts
actual_data <- data_unemployment
arima_forecast <- c(train_data, as.vector(forecast_result$mean))
naive_forecast <- c(train_data, as.vector(naive_model$mean))

# Create a data frame
df <- data.frame(
  Date = c(time(actual_data), time(actual_data), time(actual_data)),
  Value = c(as.vector(actual_data), arima_forecast, naive_forecast),
  Type = rep(c("Actual", "ARIMA", "Naive"), each = length(actual_data))
)

```

```{r}
# Plot the time series data and forecasts
p <- ggplot(df, aes(x = Date, y = Value, color = Type)) +
  geom_line(size = 1) +
  labs(title = "Unemployment Data and Forecasts",
       x = "Date",
       y = "Value",
       color = "Type") +
  theme_minimal()

# Display the plot
print(p)
```

**Overall, both models have their own strengths and weaknesses, if we only use test RMSE, we would go with ARIMA model**

**Unemployment rate is already seasonally adjusted, so we would not do SARIMA on that.**

