---
title: "Data Visualization"
author: "Yifeng Chen"
date: "2023-04-26"
output: html_document
---

```{r, echo=FALSE, include=FALSE}
library(plotly)
library(dplyr)
library(gridExtra)
library(vars)
library(GGally)
library(lmtest)
library(ggplot2)
library(tseries)
library(forecast)
```

```{r, echo=FALSE, include=FALSE}
grossDP_us <- read.csv("Gross Domestic Product US.csv")
consumerPI_us <- read.csv("Consumer Price Index US.csv")
realDI_us <- read.csv("Real Disposable Personal Income US.csv")
unemploymentR_us <- read.csv("Unemployment Rate US.csv")
```

```{r, echo=FALSE, include=FALSE}
grossDP_us$DATE <- as.Date(grossDP_us$DATE)
consumerPI_us$DATE <- as.Date(consumerPI_us$DATE)
realDI_us$DATE <- as.Date(realDI_us$DATE)
unemploymentR_us$DATE <- as.Date(unemploymentR_us$DATE)
```

```{r, echo=FALSE, include=FALSE}
colnames(grossDP_us)[2] <- "DATA"
colnames(consumerPI_us)[2] <- "DATA"
colnames(realDI_us)[2] <- "DATA"
colnames(unemploymentR_us)[2] <- "DATA"
```

```{r, echo=FALSE, include=FALSE}
consumerPI_us <- consumerPI_us %>%
  mutate(RATE = (DATA - lag(DATA)) / lag(DATA) * 100)
```

```{r, echo=FALSE}
p <- plot_ly(data = consumerPI_us, x = ~DATE, y = ~RATE, type = 'scatter', mode = 'lines+markers', name = 'ConsumerPIR_us') %>%
  add_trace(data = unemploymentR_us, x = ~DATE, y = ~DATA, mode = 'lines+markers', name = 'UnemploymentR_us') %>%
  layout(title = 'Trends in Consumer Price Index Rate and Unemployment Rate in the US',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Value'))

p
```

**Base on the behavior, the consumer price index rate and the unemployment rate seem to be additive.**

```{r, echo=FALSE}
p <- plot_ly(data = grossDP_us, x = ~DATE, y = ~DATA, type = 'scatter', mode = 'lines+markers', name = 'grossDP_us') %>%
  add_trace(data = realDI_us, x = ~DATE, y = ~DATA, mode = 'lines+markers', name = 'realDI_us') %>%
  add_trace(data = consumerPI_us, x = ~DATE, y = ~DATA, mode = 'lines+markers', name = 'consumerPI_us') %>%
  layout(title = 'Trends in Gross Domestic Product, Real Disposable Income, and Consumer Price Index in the US',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Value'))

p
```

**Base on the behavior, the GDP, consumer price index and the real disposable income seem to be multiplicative.**

**There are some sharp drops around 2000, 2008, recent periods and more due to financial crisis, dot com bubble and covid but overall the rate should remain around a safe range and the GDP and the rest should continue to grow in the long run.**

```{r, echo=FALSE, include=FALSE}
merged_data <- Reduce(function(x, y) merge(x, y, by = "DATE"), list(grossDP_us, consumerPI_us, realDI_us, unemploymentR_us))
colnames(merged_data)[2] <- "grossDP_us"
colnames(merged_data)[3] <- "consumerPI_us"
colnames(merged_data)[4] <- "consumerPIR_us"
colnames(merged_data)[5] <- "realDI_us"
colnames(merged_data)[6] <- "unemploymentR_us"
```

```{r, echo=FALSE}
p <- ggpairs(merged_data, columns = c(2, 3, 4, 5, 6)) + ggtitle("Scatterplot Matrix of all variables")

p
```

```{r, echo=FALSE}
cor(merged_data[,2:6])
```

**the correlation matrix suggests that there is a strong positive linear relationship between GDP and CPI, and between GDP and RDI. It also suggests a moderate negative linear relationship between CPI and CPIR.**